Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-01-24 article/article.Rmd
+ fixes to article + fixes to stats + remove debugging that manually overrides species

## Results
<!-- Single species prioritizations -->
Surrogates-targets improved the performance of single-species prioritizations (Figure 1). There were significant differences among the performance of prioritizations depending on the targets used to generate them and if they were assessed according to adaptive or neutral genetic variation (${\chi^{2}}_{`r single.spp.AOV1[['Df']][1]`,`r single.spp.AOV1[['Df']][2]`} = `r round(single.spp.AOV1[['Chisq']][2],2)`$, $P `r pretty.pval(single.spp.AOV1[['Pr(>Chisq)']][2])`$). Specifically, the amount-based prioritization captured much less adaptive and neutral genetic variation than either the surrogate-based ($Z = `r round(single.spp.MCP$test$tstat[['Surrogate.Adaptive variation - Amount.Adaptive variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Surrogate.Adaptive variation - Amount.Adaptive variation']])`$; $Z = `r round(single.spp.MCP$test$tstat[['Surrogate.Neutral variation - Amount.Neutral variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Surrogate.Neutral variation - Amount.Neutral variation']])`$) or genetic-based ($Z = `r round(single.spp.MCP$test$tstat[['Genetic.Adaptive variation - Amount.Adaptive variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Adaptive variation - Amount.Adaptive variation']])`$; $Z = `r round(single.spp.MCP$test$tstat[['Genetic.Neutral variation - Amount.Neutral variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Neutral variation - Amount.Neutral variation']])`$) prioritizations. No differences were detected in the proportion of adaptive ($Z = `r round(single.spp.MCP$test$tstat[['Surrogate.Adaptive variation - Genetic.Adaptive variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Surrogate.Adaptive variation - Genetic.Adaptive variation']])`$) or neutral genetic variation ($Z = `r round(single.spp.MCP$test$tstat[['Surrogate.Neutral variation - Genetic.Neutral variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Surrogate.Neutral variation - Genetic.Neutral variation']])`$) secured by the the surrogate-based and genetic-based prioritizations. However, these patterns were not reflected in the multi-species prioritizations.

@@

<!-- Correlation analysis -->
The surrogate variables showed a greater utility for adaptive genetic variation than neutral variation (Figure 4). The accumulation rate ($r$) was `r round(summary(env.correlation.NLMM)$coefficients[1],2)` $\pm$ `r round(summary(env.correlation.NLMM)$coefficients[2],2)` S.E. for the proportion of adaptive variation captured in a prioritization using environmental surrogates. On the other hand, the rate ($r$) was `r round(summary(geo.correlation.NLMM)$coefficients[1],2)` $\pm$ `r round(summary(geo.correlation.NLMM)$coefficients[2],2)` S.E. for neutral variation using geographic surrogates. Thus minor increases in the target for either surrogate results in major increases in the benefit of the prioritization. However, the coefficient was much larger for the adaptative genetic variation so high targets are not required to acheive substantial gains in adaptive genetic variaition.

>>>>

## Results
<!-- Single species prioritizations -->
Surrogates-targets improved the performance of single-species prioritizations (Figure 1). There were significant differences among the performance of prioritizations depending on the targets used to generate them and if they were assessed according to adaptive or neutral genetic variation (${\chi^{2}}_{`r single.spp.AOV1[['Df']][1]`,`r single.spp.AOV1[['Df']][2]`} = `r round(single.spp.AOV1[['Chisq']][2],2)`$, $P `r pretty.pval(single.spp.AOV1[['Pr(>Chisq)']][2])`$). Specifically, the amount-based prioritization captured much less adaptive and neutral genetic variation than either the surrogate-based ($Z = `r round(single.spp.MCP$test$tstat[['Surrogate.Adaptive variation - Amount.Adaptive variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Surrogate.Adaptive variation - Amount.Adaptive variation']])`$; $Z = `r round(single.spp.MCP$test$tstat[['Surrogate.Neutral variation - Amount.Neutral variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Surrogate.Neutral variation - Amount.Neutral variation']])`$) or genetic-based ($Z = `r round(single.spp.MCP$test$tstat[['Genetic.Adaptive variation - Amount.Adaptive variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Adaptive variation - Amount.Adaptive variation']])`$; $Z = `r round(single.spp.MCP$test$tstat[['Genetic.Neutral variation - Amount.Neutral variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Neutral variation - Amount.Neutral variation']])`$) prioritizations. No differences were detected in the proportion of adaptive ($Z = `r round(single.spp.MCP$test$tstat[['Genetic.Adaptive variation - Surrogate.Adaptive variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Adaptive variation - Surrogate.Adaptive variation']])`$) or neutral genetic variation ($Z = `r round(single.spp.MCP$test$tstat[['Genetic.Neutral variation - Surrogate.Neutral variation']],2)`$, $P `r pretty.pval(single.spp.MCP$test$pvalue[['Genetic.Neutral variation - Surrogate.Neutral variation']])`$) secured by the the surrogate-based and genetic-based prioritizations. However, these patterns were not reflected in the multi-species prioritizations.

@@

<!-- Correlation analysis -->
The surrogate variables showed a greater utility for neutral genetic variation than adaptive variation (Figure 4). The accumulation rate ($r$) was `r round(summary(env.correlation.NLMM)$coefficients[1],2)` $\pm$ `r round(summary(env.correlation.NLMM)$coefficients[2],2)` S.E. for the proportion of adaptive variation captured in a prioritization using environmental surrogates. On the other hand, the rate ($r$) was `r round(summary(geo.correlation.NLMM)$coefficients[1],2)` $\pm$ `r round(summary(geo.correlation.NLMM)$coefficients[2],2)` S.E. for neutral variation using geographic surrogates. Thus minor increases in the target for either surrogate results in major increases in the benefit of the prioritization. However, the coefficient was much larger for neutral genetic variation so minor increases in the target can acheive substantial gains.

### 2016-01-24 article/figures.Rmd
+ fixes to article + fixes to stats + remove debugging that manually overrides species

```{r, echo=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of single species prioritizations. Single species prioritizations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Violin plots show the performance of prioritizations generated using these three sets of targets. Points denote the values associated with each prioritization. Letters denote significant differences ($P < 0.05$).'}
# prepare data for plotting
single.spp.PDF <- expand.grid(
	Prioritisation=unique(single.spp.SDF$Prioritisation),
	Metric=unique(single.spp.SDF$Metric)
) %>% mutate(
	letters=toupper(cld(single.spp.MCP)$mcletters$Letters),
	letter.height=max(single.spp.SDF$value)+(diff(range(single.spp.SDF$value))*0.1)
)
# make plot
ggplot(aes(x=Metric,y=value,fill=Prioritisation),
	data=single.spp.SDF) +
	geom_violin(position=position_dodge(0.9)) +
	geom_point(alpha=0.2,position=position_jitterdodge(dodge.width=0.9, jitter.width=0.25)) +
	geom_text(aes(x=Metric, y=letter.height, label=letters),
		data=single.spp.PDF, position=position_dodge(0.9)) +
	scale_fill_manual(name='Prioritisation',
		values=c('grey80','grey50','grey20')) +
	ylab('Genetic variation secured (%)') +
	xlab('') +
	theme_classic()
```

@@

```{r, echo=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of multi-species prioritizations. Three prioritizations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Data shows the performance of these prioritizations based on how much genetic variation they explain. See Figure 1 caption for conventions.'}
# prepare data for plotting
multi.spp.PDF <- expand.grid(
	Prioritisation=unique(multi.spp.SDF$Prioritisation),
	Metric=unique(multi.spp.SDF$Metric)
) %>% mutate(
	letters=toupper(cld(multi.spp.MCP)$mcletters$Letters),
	letter.height=max(multi.spp.SDF$value)+(diff(range(multi.spp.SDF$value))*0.1)
)
# make plot
ggplot(aes(x=Metric,y=value,fill=Prioritisation),
	data=multi.spp.SDF) +
	geom_violin(position=position_dodge(0.9)) +
	geom_point(alpha=0.2,position=position_jitterdodge(dodge.width=0.9, jitter.width=0.25)) +
	geom_text(aes(x=Metric, y=letter.height,label=letters),
		data=multi.spp.PDF, position=position_dodge(0.9)) +
	scale_fill_manual(name='Prioritisation',
		values=c('grey80','grey50','grey20')) +
	ylab('Genetic variation secured (%)') +
	xlab('') +
	theme_classic()
```

```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations. Each panel shows the relationship for a different type of genetic variation. Each line denotes a different species. Lines are transparent and so darker regions show areas where species have similar relationships.'}
# make plots
p1 <- ggplot(data=env.pareto.DF) +
	geom_line(aes(x=Surrogate.target,y=adaptive.held,group=Species),
		alpha=0.3) +
	xlab('Environmental variation secured (%)') +
	ylab('Adaptive genetic\nvariation secured (%)') +
	theme_classic()
p2 <- ggplot(data=geo.pareto.DF) +
	geom_line(aes(x=Surrogate.target,y=neutral.held,group=Species),
		alpha=0.3) +
	xlab('Geographic variation secured (%)') +
	ylab('Neutral genetic\nvariation secured (%)') +
	theme_classic()
grid.arrange(p1, p2, nrow=1)
```

>>>>

```{r, echo=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of single species prioritizations. Single species prioritizations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Violin plots show the performance of prioritizations generated using these three sets of targets. Letters denote significant differences ($P < 0.05$).'}
# prepare data for plotting
single.spp.PDF <- expand.grid(
	Prioritisation=unique(single.spp.SDF$Prioritisation),
	Metric=unique(single.spp.SDF$Metric)
) %>% mutate(
	letters=toupper(cld(single.spp.MCP)$mcletters$Letters),
	letter.height=max(single.spp.SDF$value, na.rm=TRUE)+(diff(range(single.spp.SDF$value, na.rm=TRUE))*0.1)
)
# make plot
suppressWarnings({
	ggplot(aes(x=Metric,y=value,fill=Prioritisation),
		data=single.spp.SDF) +
		geom_violin(position=position_dodge(0.9)) +
		geom_text(aes(x=Metric, y=letter.height, label=letters),
			data=single.spp.PDF, position=position_dodge(0.9)) +
		scale_fill_manual(name='Prioritisation',
			values=c('grey80','grey50','grey20')) +
		ylab('Genetic variation secured (%)') +
		xlab('') +
		theme_classic()
})
```

@@

```{r, echo=FALSE, fig.width=6, fig.height=3, fig.cap='Summary of multi-species prioritizations. Three prioritizations were generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Data shows the performance of these prioritizations based on how much genetic variation they explain. See Figure 1 caption for conventions.'}
# prepare data for plotting
multi.spp.PDF <- expand.grid(
	Prioritisation=unique(multi.spp.SDF$Prioritisation),
	Metric=unique(multi.spp.SDF$Metric)
) %>% mutate(
	letters=toupper(cld(multi.spp.MCP)$mcletters$Letters),
	letter.height=max(multi.spp.SDF$value, na.rm=TRUE)+(diff(range(multi.spp.SDF$value, na.rm=TRUE))*0.1)
)
# make plot
suppressWarnings({
	ggplot(aes(x=Metric,y=value,fill=Prioritisation),
		data=multi.spp.SDF) +
		geom_violin(position=position_dodge(0.9)) +
		geom_text(aes(x=Metric, y=letter.height,label=letters),
			data=multi.spp.PDF, position=position_dodge(0.9)) +
		scale_fill_manual(name='Prioritisation',
			values=c('grey80','grey50','grey20')) +
		ylab('Genetic variation secured (%)') +
		xlab('') +
		theme_classic()
})
```

```{r, echo=FALSE, fig.width=8, fig.height=3, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations. The left panel shows the relationship between environmental surrogates and adaptive genetic variation, and the right for geographic surrogates and neutral genetic variation. Transparent gray lines denote trneds for each species, and black lines show general trends across all species.'}
### prepare data frames for plotting
## environmental vs adaptive
# random effects
env.correlation.RPDF <- expand.grid(Surrogate.target=seq(min(env.correlation.DF$Surrogate.target), max(env.correlation.DF$Surrogate.target), length.out=100),
	Species=unique(env.correlation.DF$Species))
env.correlation.RPDF$r.ranef <- fixef(env.correlation.NLMM)[[1]] + ranef(env.correlation.NLMM)$Species$r[env.correlation.RPDF$Species]
env.correlation.RPDF$fit <- 1 - (1 - 0) * exp(- env.correlation.RPDF$r.ranef * env.correlation.RPDF$Surrogate.target)
# fixed effects
env.correlation.FPDF <- data.frame(Surrogate.target=seq(min(env.correlation.DF$Surrogate.target), max(env.correlation.DF$Surrogate.target), length.out=100))
env.correlation.FPDF$fit <- 1 - (1 - 0) * exp(- fixef(env.correlation.NLMM)[[1]] * env.correlation.FPDF$Surrogate.target)

## geographic vs neutral
# random effects
geo.correlation.RPDF <- expand.grid(Surrogate.target=seq(min(geo.correlation.DF$Surrogate.target), max(geo.correlation.DF$Surrogate.target), length.out=100),
	Species=unique(geo.correlation.DF$Species))
geo.correlation.RPDF$r.ranef <- fixef(geo.correlation.NLMM)[[1]] + ranef(geo.correlation.NLMM)$Species$r[geo.correlation.RPDF$Species]
geo.correlation.RPDF$fit <- 1 - (1 - 0) * exp(- geo.correlation.RPDF$r.ranef * geo.correlation.RPDF$Surrogate.target)
# fixed effects
geo.correlation.FPDF <- data.frame(Surrogate.target=seq(min(geo.correlation.DF$Surrogate.target), max(geo.correlation.DF$Surrogate.target), length.out=100))
geo.correlation.FPDF$fit <- 1 - (1 - 0) * exp(- fixef(geo.correlation.NLMM)[[1]] * geo.correlation.FPDF$Surrogate.target)

### make plots
suppressWarnings({
	p1 <- ggplot() +
		geom_point(aes(x=Surrogate.target, y=adaptive.held), data=env.correlation.DF) +
		geom_line(aes(x=Surrogate.target,y=fit,group=Species), data=env.correlation.RPDF, alpha=0.3) +
		geom_line(aes(x=Surrogate.target,y=fit), data=env.correlation.FPDF, size=1.5) +
		xlab('Environmental variation secured (%)') +
		ylab('Adaptive genetic\nvariation secured (%)') +
		theme_classic()
	p2 <- ggplot() +
		geom_point(aes(x=Surrogate.target, y=neutral.held), data=geo.correlation.DF) +
		geom_line(aes(x=Surrogate.target,y=fit,group=Species), data=geo.correlation.RPDF, alpha=0.3) +
		geom_line(aes(x=Surrogate.target,y=fit), data=geo.correlation.FPDF, size=1.5) +
		xlab('Geographic variation secured (%)') +
		ylab('Neutral genetic\nvariation secured (%)') +
		theme_classic()
	grid.arrange(p1, p2, nrow=1)
})
```

### 2016-01-24 article/supporting_information.Rmd
+ fixes to article + fixes to stats + remove debugging that manually overrides species

```{r, echo=FALSE, fig.width=6.5, fig.height=6.5, fig.cap='Distribution of populations. Each panel depicts a different species. Squares denote planning units. Points denote the location of sampled individuals, and their colors denote the distribution of populations. Note that points have been slightly jittered so that individuals in close proximity are visible.'}
# define function
plotLST <- llply(
	seq_along(unique(spp.samples.DF$species)),
	.fun=function(i) {
		# init
		curr.pts <- filter(
			spp.sample.PTS@data,
			species==unique(spp.samples.DF$species)[i]
		) %>% mutate(
			population=spp.Adegenet.LST[[i]]$grp
		)
		# create plot
		ggplot() +
			geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
				fill='grey20', color='grey80') +
			geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
				group='group'), fill='transparent', color='grey10') +
			geom_point(data=curr.pts, aes_string(x='sample.longitude', y='sample.latitude', color='population'),
				alpha=0.3, size=3, position='jitter') +
			guides(fill=guide_legend(title=' ')) +
			theme_classic() +
			theme(axis.ticks=element_blank(), axis.text=element_blank(),
				plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
			coord_cartesian(
				xlim=buffered.range(grid.FPLY$long, 0.05),
				ylim=buffered.range(grid.FPLY$lat, 0.05)
			) +
			xlab('') +
			ylab('') +
			ggtitle(gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE))
	}
)
# render plot
do.call(
	grid.arrange,
	append(
		plotLST,
		list(ncol=6)
	)
)
```

@@

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}. See Figure S6 caption for conventions.')}
if (23 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(23)
```

>>>>

```{r, echo=FALSE, fig.width=6.5, fig.height=6.5, fig.cap='Distribution of populations. Each panel depicts a different species. Squares denote planning units. Points denote the location of sampled individuals, and their colors denote the distribution of populations. Note that points have been slightly jittered so that individuals in close proximity are visible.'}
# define function
plotLST <- llply(
	# seq_along(unique(spp.samples.DF$species)),
	seq_along(unique(spp.samples.DF$species)[-23]), # omit spp 23 from current version
	.fun=function(i) {
		# init
		curr.pts <- filter(
			spp.sample.PTS@data,
			species==unique(spp.samples.DF$species)[i]
		) %>% mutate(
			population=spp.Adegenet.LST[[i]]$grp
		)
		# create plot
		ggplot() +
			geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
				fill='grey20', color='grey80') +
			geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
				group='group'), fill='transparent', color='grey10') +
			geom_point(data=curr.pts, aes_string(x='sample.longitude', y='sample.latitude', color='population'),
				alpha=0.3, size=3, position='jitter') +
			guides(fill=guide_legend(title=' ')) +
			theme_classic() +
			theme(axis.ticks=element_blank(), axis.text=element_blank(),
				plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
			coord_cartesian(
				xlim=buffered.range(grid.FPLY$long, 0.05),
				ylim=buffered.range(grid.FPLY$lat, 0.05)
			) +
			xlab('') +
			ylab('') +
			ggtitle(gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE))
	}
)
# render plot
do.call(
	grid.arrange,
	append(
		plotLST,
		list(ncol=6)
	)
)
```

@@

```{r, echo=FALSE, fig.width=6.5, fig.height=3.25, fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}. See Figure S6 caption for conventions.')}
# if (23 <= general.params.LST[[MODE]]$n.spp) plot.spp.mds(23)
```
